Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-20 Writeup/ms.Rmd
modif to manuscript

 Identifying a spoken word in a referential context requires both the ability to process and integrate multimodal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model of the task which provides an account of how auditory and visual cues are combined optimally. Model predictions are tested in three experiments where word recognition is made under two kinds of uncertainty: category ambiguity and/or distorting noise. In all cases, the optimal model explains much of the variance in human mean judgments. In particular, when the signal is not distorted with noise, participants weight the auditory and visual cues optimally, that is, according to the relative reliability of each modality. But when one modality has noise added to it, human perceivers systematically prefer the unperturbed modality to a greater extent than the optimal model does. The study provides a formal framework which helps us understand precisely how word form and word meaning interact in word recognition under uncertainty. Moreover it offers a first step towards a model that accounts for form-meaning synergy in early word learning.

@@

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
library(ggthemes)
#theme_set(theme_few())
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

```{r}
summarise <- dplyr::summarise
rename <- dplyr::rename
```

@@

#Summary
sounds_exp1 <- exp1_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

@@

exp_uni_data <- bind_rows(sounds_exp1, concepts_exp1,
                     sounds_exp2, concepts_exp2,
                     sounds_exp3, concepts_exp3)

@@

Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell "bee!" at a picnic, as a honey bee buzzes around the food. Identifying a word involves processing the auditory information as well as other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm). A word is successfully identified when information from these modalities provide convergent evidence. However, word identification takes place in a noisy world, and the cues received through each modality may not provide a definitive answer. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically [@hillenbrand1995]. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling "pea" and not "bee". Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the insect was a beetle or a fly instead.

How does the listener deal with uncertainty to identify the speaker's intended word? One rigorous way to approach this question is through conducting an *ideal observer* analysis. This research strategy provides a characterization of the task/goal and shows what the optimal performance should be under this characterization.^[It is, thus, a general instance of the rational approach to cognition [@anderson90]. It can also be seens as an instance of Marr's computational level of analysis.] When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. When the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is supposed to be used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003; @rahnev2018]. When there is a deviation from the ideal, it can reveal extra constraints on human cognition, such as limitations on the working memory or attentional resources.

The ideal observer analysis has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]. In particular, using this research strategy, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (VOT) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, which is a phenomenon that involves reduced discriminability near prototypical sounds in the native language [@kuhl1991]. They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty. Both @clayard08 and @feldman2009 explored optimal performance under uncertainty in the auditory modality. There is, however, extensive evidence that information from the visual modality, such as the speaker's facial features,  also influences speech understanding [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how probabilistic cues from speech and lip movements can be optimally combined. They showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer.

This previous research, however, did not systematically study speech understanding when the visual information is obtained, not through the speaker's facial features, but through the referential context. In fact, experimental findings showed that information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the listener has to combine cues from the sound and the referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does it make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

On the face of it, the question of combining information from the sound and the visual referent might seem similar to that of audio-visual speech integration. Nevertheless, there are at least two fundamental differences between these two cases, and both can influence the way the auditory and visual cues are combined:

\noindent First, in the case of audio-visual speech, both modalities offer information about the same underlying speech category. They may differ only in terms of their informational reliability. In a referential context, however, the auditory and visual modalities are additionally different in terms of the roles they play in the referential process: the auditory input represents the *symbol* whereas the visual input represents the *meaning*. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].^[There is a debate as to whether speech is privileged for children and adults for the same reasons. Whereas some researchers suggest that speech is privileged for both children and adults because of its ability to refer [e.g., @waxman1995], others suggest that speech might *not* have a referential status from the start. Rather, speech might be prefered by children only because of a low level auditory ``overshadowing'' [e.g., @sloutsky2003].] Thus, in a referential context, it is possible that listeners do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a sub-optimal bias for the auditory modality beyond what is expected from informational reliability alone.

\noindent Second in the case of audio-visual speech, the auditory and visual stimuli are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between the auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. In sum, cue combination in the case of arbitrary audio-visual associations (word-referent) is likely to be less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality.  We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model (which is fit to human responses). Comparing parameter estimates between these two formulations allows us to quantify the degree of deviation from optimality.

@@

We construct an ideal observer model that combines probabilistic information from the auditory and visual modalities. In contrast to the  model used in most research on multisensory integration [e.g., @ernst02]---which typically studies continuous stimuli (e.g., size, location)---this probabilistic information in our case cannot be characterized with *sensory noise/variability*, only.  Indeed, our task involves responses over categorical variables (phonemes and concepts), and therefore, the optimal model in our case should take into account, not only the noise variability around an individual perceptual estimate, but also its *categorical variability*, i.e., the uncertainty related to whether this perceptual estimate belongs to a given category [see also @Bankieris17; @bejjanki2011]. In what follows, we describe a model that accounts for both type of variability. First, we describe the model in the simplified case of categorical variability only. Second, we augment this simplified model to account for sensory noise.

@@

\noindent This assumption simply says that, given a word-object mapping, e.g., $W=$("cat"-CAT), variation in the way "cat" is pronounced does not correlate with changes in any visual property of the object CAT, which is a valid assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

So far, we only accounted for categorical variability. For instance, if the speaker generates a target production $a_t$ from an auditory category
$p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})$, the ideal model assumes that it has direct access to this production token (i.e., $a=a_t$), and that all uncertainty is about the category membership of this token. However, we might also want to account for noise in the brain and/or in the environment.  For example, the observer might not have access to the exact produced target, but only to the target perturbed by noise. If we assume this noise to be normally distributed, that is,  $p(a | a_t) \sim N(a_t, \sigma^2_{N_A})$, then integrating over $a_t$ leads to the following simple expression:
$$ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})$$
Similarly, in the case of sensory noise in the visual modality, we get
$$ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})$$
Finally, using exactly the same derivation as above, we end up with the following multimodal weighting scheme in the optimal combination model (Equation 1) which takes into account both categorical and sensory variability:

@@

The predictions of the optimal model will be compared to two baselines. The first baseline is a visual model which assumes that participants rely only on visual information, and an auditory model, which assumes that participants rely only on auditory information. More precisely, these baseline models assume that the participants' responses in the bimodal condition will not be different from their response in either the visual-only or the auditory-only condition. If, as we expect, the participants rely on both the auditory and the visual modalities to make decision in the bimodal condition, the optimal model would explain more variance in human responses than the visual or the auditory model do.

@@

We compared these models to human responses in three experiments. In Experiment 1, we studied the case where bimodal uncertainty was due to ambiguity in terms of category membership, without any additional background noise. In Experiment 2 and 3 we added background noise on top of ambiguity in category membership.

@@

In this Experiment, we start with testing the predictions in the case where uncertainty is due to categorical variability only (i.e., ambiguity in terms of category memebership). We do not add any external noise to the background and we assume that internal sensory noise is negligible compared to categorical variability ($\sigma^2_{A} \gg \sigma^2_{N_A}$, and $\sigma^2_{V} \gg \sigma^2_{N_V}$). Thus, we use the following cue weighting scheme:

@@

@@

###### Extract values from saved data

@@

#Fit the non-linear function

##Experiment 1
#############
sound_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrA)*sound_dist+(8/vrA))), data=sound_all_exp1, start = list(e=0, vrA=2))

concept_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrV)*concept_dist+(8/vrV))), data=concept_all_exp1, start = list(e=0, vrV=2))

#Here I should test the cross-validation

joint_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrA_j)*sound_dist+(-4/vrV_j)*concept_dist+(8/vrA_j)+(8/vrV_j))), data=joint_all_exp1, start = list(e=0, vrA_j=2, vrV_j=2))

@@

x <- seq(0, 4, 0.01)

@@

all_uni_model <- bind_rows(uniS_nl1, uniV_nl1,
                           uniS_nl2, uniV_nl2,
                           uniS_nl3, uniV_nl3)

all_plot <- ggplot(exp_uni_data, 
       aes(x = distance, y = mean)) + 
  geom_point()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  position = position_dodge(width = .1)) + 
  #geom_line(data=uniV,aes(x=xV, y=yV))+
  geom_line(data=all_uni_model, aes(x=distance, y=prediction), col='black')+
  xlab("Distance") +ylab("Prob. different")+
  scale_y_continuous(limits = c(0, 1))+#theme(aspect.ratio = 0.7)+
  theme_few()+
  theme(aspect.ratio = 0.7) + facet_grid(Experiment ~ Condition)
  #stat_function(fun = Logistic_v, colour="red"))

@@

#### Bimodal condition
In this condition, only the descriptive model is fit to the data, using the expression of the posterior (Equation 1). Since the values of the means are fixed, we have 3 free parameters: the variances for the visual and the auditory modalities, respectively, and $b$, the response bias.  The visual and auditory baselines as well as the optimal model are not fit to bimodal data, but their predictions are tested against these bimodal data. All these normative models use the variances derived from the unimodal data and the bias term derived from the bimodal data.

Although the paradigm is within-subjects, we did not have enough statistical power to fit a different model for each individual participant. Instead, models were constructed with data collapsed across all participants. That being said, the distribution of responses from individual participants will also be studied. The fit was done with a nonlinear least squares regression using the NLS package in R [@bates88]. We computed the values of the parameters, as well as their 95% confidence intervals, through non-parametric bootstrap (using 10000 iterations).

@@

#All models

#Here we use the values obtained from the fitting using bootstrap sampling to

#The fit
model_fit0 <- function (x,y) {
   1/(1 + exp((-4/vrJ_A_nl1)*x+(-4/vrJ_V_nl1)*y+(8/vrJ_A_nl1)+(8/vrJ_V_nl1)))
}

@@

#The fit
model_fit1 <- function (x,y) {
   1/(1 + (1-eJ_nl1)*exp((-4/vrJ_A_nl1)*x+(-4/vrJ_V_nl1)*y+(8/vrJ_A_nl1)+(8/vrJ_V_nl1)))
}

#The ideal
model_ideal1 <- function (x,y) {
    1/(1 + (1-eJ_nl1)*exp((-4/vrA_nl1)*x+(-4/vrV_nl1)*y+(8/vrA_nl1)+(8/vrV_nl1)))
}

@@

#The fit
model_fit2 <- function (x,y) {
   1/(1 + (1-eJ_nl2)*exp((-4/vrJ_A_nl2)*x+(-4/vrJ_V_nl2)*y+(8/vrJ_A_nl2)+(8/vrJ_V_nl2)))
}

#The ideal
model_ideal2 <- function (x,y) {
    1/(1 + (1-eJ_nl2)*exp((-4/vrA_nl2)*x+(-4/vrV_nl2)*y+(8/vrA_nl2)+(8/vrV_nl2)))
}

@@

#The fit
model_fit3 <- function (x,y) {
   1/(1 + (1-eJ_nl3)*exp((-4/vrJ_A_nl3)*x+(-4/vrJ_V_nl3)*y+(8/vrJ_A_nl3)+(8/vrJ_V_nl3)))
}

#The ideal
model_ideal3 <- function (x,y) {
    1/(1 + (1-eJ_nl3)*exp((-4/vrA_nl3)*x+(-4/vrV_nl3)*y+(8/vrA_nl3)+(8/vrV_nl3)))
}

@@

@@

@@

correlation_plot <- ggplot(models_all, 
       aes(x = pred, y = joint, col = factor(concept_dist), 
           shape = factor(sound_dist))) + 
  geom_point()+
 #geom_pointrange(aes(ymin = summary_ci_lower, ymax = summary_ci_upper), 
  #                position = position_dodge(width = .1), size=0.2) +

@@

#Explore cross-validation here. As the numbers show, the predictions on the held-out subset is almost as acurate as the predcitions on the training set.  Over-fitting is generally not a problem with such simple models)

#Here is use hald the responses to predic the other half
#The function takes the data to be cross-validated and return a score (coefficient of determination)

@@

crossValid_exp1 <- replicate(10, {
  crossValid(joint_all_exp1)
  })

@@

crossValid_exp2 <- replicate(10, {
  crossValid(joint_all_exp2)
  })

@@

crossValid_exp3 <- replicate(10, {
  crossValid(joint_all_exp3)
  })

@@

  #I need to take the average of this over multiple runs

@@

#Here do some quatiative measure:

>>>>

 Identifying a spoken word in a referential context requires both the ability to process and integrate multimodal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model which provides an account of how auditory and visual cues are combined optimally. Model predictions are tested in three experiments where word recognition is made under two kinds of uncertainty: category ambiguity and/or distorting noise. In all cases, the optimal model explains much of the variance in human mean judgments. In particular, when the signal is not distorted with noise, participants weight the auditory and visual cues optimally, that is, according to the relative reliability of each modality. But when one modality has noise added to it, human perceivers systematically prefer the unperturbed modality to a greater extent than the optimal model does. The study provides a formal framework which helps us understand precisely how word form and word meaning interact in word recognition under uncertainty. Moreover it offers a first step towards a model that accounts for form-meaning synergy in early word learning.

@@

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
library(ggthemes)
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

@@

#Summary

sounds_exp1 <- exp1_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

@@

#Combine all
exp_uni_data <- bind_rows(sounds_exp1, concepts_exp1,
                     sounds_exp2, concepts_exp2,
                     sounds_exp3, concepts_exp3)

@@

Language uses symbols expressed in one modality, e.g., the auditory modality, in the case of speech, to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell "bee!" at a picnic, as a honey bee buzzes around the food. Identifying a word involves processing the auditory information as well as other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm). A word is successfully identified when information from these modalities provide convergent evidence. However, word identification takes place in a noisy world, and the cues received through each modality may not provide a definitive answer. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically [@hillenbrand1995]. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling "pea" and not "bee". Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the insect was a beetle or a fly instead.  How does the listener deal with such multimodal uncertainty to recognize the speaker's intended word?

The task of matching the sound to the corresponding visual object has been extensively studied in the developmental literature since it is considered to be an crucial instance of early word learning. For example, many studies focused on how children succeed in this task despite high referential ambiguity [@smith08; @pinker1989; @medina2011; @vouloumanos2008; @yurovsky2015; @suanda2014; @vlach2013]. However, even when they know the exact meanings of the words, listeners (both children and adults) often face the task of recognizing which word the speaker has uttered, especially under noisy circumstances. The purpose of the current study is to explore the special case of word recognition under uncertainty when adult observers have access to multimodal cues from the speech and the referent.

One rigorous way to approach this question is through conducting an *ideal observer* analysis. This research strategy provides a characterization of the task/goal and shows what the optimal performance should be under this characterization.^[It is, thus, a general instance of the rational approach to cognition [@anderson90]. It can also be seen as an instance of Marr's computational level of analysis.] When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. The ideal observer model can be seen as a theoretical upper limit on performance. It is not so much a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003; @rahnev2018]. When there is a deviation from the ideal, it can reveal extra constraints on human cognition, such as limitations on the working memory or attentional resources.  The ideal observer analysis has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]

Some of these ideal-observer-based studies are closely related to the question we are addressing in the current work. For instance, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (Voice Onset Time) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, which is a phenomenon that involves reduced discriminability near prototypical sounds in the native language [@kuhl1991]. They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty.

Besides the acoustic cues explored in @clayard08 and @feldman2009, there is extensive evidence that information from the visual modality, such as the speaker's facial features,  also influences speech understanding [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how probabilistic cues from speech and lip movements can be optimally combined. They showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer. This previous research, however, did not systematically study speech understanding when the visual information is obtained, not through the speaker's facial features---as in audio-visual speech perception, but through the referential context. In fact, experimental findings showed that information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the listener has to combine cues from the sound and the referent.

On the face of it, the question of combining information from the sound and the visual referent might seem similar to that of audio-visual speech integration. Nevertheless, there are at least two fundamental differences between these two cases, and both can influence the way the auditory and visual cues are combined: First, in the case of audio-visual speech, both modalities offer information about the same underlying speech category. They may differ only in terms of their informational reliability. In a referential context, however, the auditory and visual modalities play different roles in the referential process---in addition to possible differences in informational reliability. Indeed, the auditory input represents the *symbol* whereas the visual input represents the *meaning*. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].^[There is a debate as to whether speech is privileged for children and adults for the same reasons. Whereas some researchers suggest that speech is privileged for both children and adults because of its ability to refer [e.g., @waxman1995], others suggest that speech might *not* have a referential status from the start. Rather, speech might be preferred by children only because of a low level auditory ``overshadowing'' [e.g., @sloutsky2003].] Thus, in a referential context, it is possible that listeners do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a sub-optimal bias for the auditory modality beyond what is expected from informational reliability alone.

Second, in the case of audio-visual speech, the auditory and visual stimuli are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between the auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. In sum, cue combination in the case of arbitrary audio-visual associations (word-referent) is likely to be less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

## The current study

We investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does it make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model (which is fit to human responses). Comparing parameter estimates between these two formulations allows us to quantify the degree of deviation from optimality.

@@

We construct an ideal observer model that combines probabilistic information from the auditory and visual modalities. In contrast to the  model used in most research on multisensory integration [e.g., @ernst02]---which typically studies continuous stimuli (e.g., size, location)---the probabilistic information in our case cannot be characterized with *sensory noise*, only.  Indeed, our task involves responses over categorical variables (phonemes and concepts), and therefore, the optimal model should take into account, not only the noise variability around an individual perceptual estimate, but also its *categorical variability*, i.e., the uncertainty related to whether this perceptual estimate belongs to a given category [see also @Bankieris17; @bejjanki2011]. In what follows, we describe a model that accounts for both type of variability. First, we describe the model in the simplified case of categorical variability only. Second, we augment this simplified model to account for sensory noise.

@@

\noindent This assumption says that, given a word-object mapping, e.g., $W=$("cat"-CAT), variation in the way "cat" is pronounced does not correlate with changes in any visual property of the object CAT, which is a valid assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

So far, we only accounted for categorical variability. For instance, if the speaker generates a target production $a_t$ from an auditory category
$p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})$, the ideal model assumes that it has direct access to this production token (i.e., $a=a_t$), and that all uncertainty is about the category membership of this token. However, we might also want to account for internal noise in the brain and/or external noise in the environment.  For example, the observer might not have access to the exact produced target, but only to the target perturbed by noise. If we assume this noise to be normally distributed, that is,  $p(a | a_t) \sim N(a_t, \sigma^2_{N_A})$, then integrating over $a_t$ leads to the following simple expression:
$$ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})$$
Similarly, in the case of sensory noise in the visual modality, we get
$$ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})$$
Finally, using exactly the same derivation as above, we end up with the following multimodal weighting scheme in the optimal combination model (Equation 1) which takes into account both categorical and sensory variability:

@@

The predictions of the optimal model will be compared to two baselines. The first baseline is a visual model which assumes that participants rely only on visual information, and an auditory model, which assumes that participants rely only on auditory information. More precisely, these baseline models assume that the participants' responses in the bimodal condition will not be different from their response in either the visual-only or the auditory-only condition. However, if the participants rely on both the auditory and the visual modalities to make decision in the bimodal condition, the optimal model would explain more variance in human responses than the visual or the auditory model do.

@@

We compared these models to human responses in three experiments. In Experiment 1, we studied the case where bimodal uncertainty was due to categorical variability, only. In Experiment 2 and 3 we added noise in the background on top of categorical variability.

@@

In this Experiment, we test the predictions of the model in the case where uncertainty is due to categorical variability only (i.e., ambiguity in terms of category membership). We do not add any external noise to the background and we assume that internal sensory noise is negligible compared to categorical variability ($\sigma^2_{A} \gg \sigma^2_{N_A}$, and $\sigma^2_{V} \gg \sigma^2_{N_V}$). Thus, we use the following cue weighting scheme:

@@

#Bootstrap sample of parameters

@@

###### Extract values from saved data 
##To reproduce the values, please uncomment the previous code following the function "lmfit"

@@

#Fit the posteriors to the reponses

##Experiment 1
#############

#Posterior for auditory-only condition
sound_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrA)*sound_dist+(8/vrA))), data=sound_all_exp1, start = list(e=0, vrA=2))

#Posterior for visual-only condition
concept_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrV)*concept_dist+(8/vrV))), data=concept_all_exp1, start = list(e=0, vrV=2))

#Posterior for audio-visual condition (this is the descriptive model)
joint_nl1 <- nls(answer ~ 1/(1+(1-e)*exp((-4/vrA_j)*sound_dist+(-4/vrV_j)*concept_dist+(8/vrA_j)+(8/vrV_j))), data=joint_all_exp1, start = list(e=0, vrA_j=2, vrV_j=2))

@@

### the fit functions
x <- seq(0, 4, 0.01)

@@

#### Combine all the fit functions
all_uni_model <- bind_rows(uniS_nl1, uniV_nl1,
                           uniS_nl2, uniV_nl2,
                           uniS_nl3, uniV_nl3)

### plot unimodal data + fit functions 
all_plot <- ggplot(exp_uni_data, 
       aes(x = distance, y = mean)) + 
  geom_point()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  position = position_dodge(width = .1)) + 
  #geom_line(data=uniV,aes(x=xV, y=yV))+
  geom_line(data=all_uni_model, aes(x=distance, y=prediction), col='black')+
  xlab("Distance") +ylab("Prob. different")+
  scale_y_continuous(limits = c(0, 1))+#theme(aspect.ratio = 0.7)+
  theme_few()+
  theme(aspect.ratio = 0.7) + facet_grid(Experiment ~ Condition)
  #stat_function(fun = Logistic_v, colour="red"))

@@

#### Bimodal condition
In this condition, only the descriptive model is fit to the data, using the expression of the posterior (Equation 1). Since the values of the means are fixed, we have 3 free parameters: the variances for the visual and the auditory modalities, respectively, and $b$, the response bias.  The visual and auditory baselines as well as the optimal model are not fit to the bimodal data, but their predictions are tested against these bimodal data. All these normative models use the variances derived from the unimodal data and the bias term derived from the fit to bimodal data.

Although the paradigm is within-subjects, we did not have enough statistical power to fit a different model for each individual participant. Instead, models were constructed with data collapsed across all participants. That being said, the distribution of responses from individual participants will also be analyzed. The fit was done with a nonlinear least squares regression using the NLS package in R [@bates88]. We computed the values of the parameters, as well as their 95% confidence intervals, through non-parametric bootstrap (using 10000 iterations).

@@

###Models to be tested against responses in the bimodal condition

#Descriptive model without a terms for response bias 
model_fit0 <- function (x,y) {
   1/(1 + exp((-4/vrJ_A_nl1)*x+(-4/vrJ_V_nl1)*y+(8/vrJ_A_nl1)+(8/vrJ_V_nl1)))
}

@@

#Descriptive model (parameters are derived from fit to bimodal data)
model_fit1 <- function (x,y) {
   1/(1 + (1-eJ_nl1)*exp((-4/vrJ_A_nl1)*x+(-4/vrJ_V_nl1)*y+(8/vrJ_A_nl1)+(8/vrJ_V_nl1)))
}

#The optimal model (parameters are derived from fit to unimodal data)
model_ideal1 <- function (x,y) {
    1/(1 + (1-eJ_nl1)*exp((-4/vrA_nl1)*x+(-4/vrV_nl1)*y+(8/vrA_nl1)+(8/vrV_nl1)))
}

@@

#Descriptive model (parameters are derived from fit to bimodal data)
model_fit2 <- function (x,y) {
   1/(1 + (1-eJ_nl2)*exp((-4/vrJ_A_nl2)*x+(-4/vrJ_V_nl2)*y+(8/vrJ_A_nl2)+(8/vrJ_V_nl2)))
}

#The optimal model (parameters are derived from fit to unimodal data)
model_ideal2 <- function (x,y) {
    1/(1 + (1-eJ_nl2)*exp((-4/vrA_nl2)*x+(-4/vrV_nl2)*y+(8/vrA_nl2)+(8/vrV_nl2)))
}

@@

#Descriptive model (parameters are derived from fit to bimodal data)
model_fit3 <- function (x,y) {
   1/(1 + (1-eJ_nl3)*exp((-4/vrJ_A_nl3)*x+(-4/vrJ_V_nl3)*y+(8/vrJ_A_nl3)+(8/vrJ_V_nl3)))
}

#The optimal model (parameters are derived from fit to unimodal data)
model_ideal3 <- function (x,y) {
    1/(1 + (1-eJ_nl3)*exp((-4/vrA_nl3)*x+(-4/vrV_nl3)*y+(8/vrA_nl3)+(8/vrV_nl3)))
}

@@

#Generate prediction in the bimodal condition

@@

##Combine all predictions

@@

correlation_plot <- ggplot(models_all, 
       aes(x = pred, y = joint, col = factor(concept_dist), 
           shape = factor(sound_dist))) + 
  geom_point()+
  scale_colour_solarized()+
 #geom_pointrange(aes(ymin = summary_ci_lower, ymax = summary_ci_upper), 
  #                position = position_dodge(width = .1), size=0.2) +

@@

#Here we conduct cross-validation for the descriptive model, since this model is trained and tested on the same data

#As the numbers show, the predictions on the held-out subset is almost as accurate as the predictions on the training set.  Over-fitting is generally not a problem with such simple models)

#We conduct a cross-validation where half the responses are used predict the other half

#The function "crossValid", below, takes the data to be cross-validated and return a score (coefficient of determination)

@@

#Average across N partions, to render the pdf I am using a small value, but higher values can be used for the final version
N_partitions = 10

crossValid_exp1 <- replicate(N_partitions, {
  crossValid(joint_all_exp1)
  })

@@

crossValid_exp2 <- replicate(N_partitions, {
  crossValid(joint_all_exp2)
  })

@@

crossValid_exp3 <- replicate(N_partitions, {
  crossValid(joint_all_exp3)
  })

@@

#Process and combine Real and Simulated data

@@
